---
title: "R Notebook"
output: html_notebook
---

```{r}
library(caret)
library(dplyr)
library(ggplot2)
library(tidyr)
library(xts)
library(reshape)
library(zoo)
library(urca)
library(vars)
```

```{r}
library(readxl)
library(ggplot2)
library(lattice)
library(tidyverse)
library(lubridate)
library(stringr)
library(xts)
library(urca)
library(dynlm)
library(tseries)
library(quantmod)
library(lmtest)
library(sandwich)
```



```{r}
gt <- read.csv("C:\\Users\\jacob\\OneDrive\\Documents\\EECS448\\Project\\Data\\gt_data.csv")
gt$start <- as.Date(gt$start)
head(gt)
```

```{r}
# Modify the dates so that the first Google Trend result of every month is on the first. This is so we can merge it with CPI. Google Trends was an average over a period of a week, so changing the reference date by < 7 days shouldn't impact results

gt <- gt %>%
  arrange(start) %>% 
  group_by(year = format(start, "%Y"), month = format(start, "%m")) %>% # Group by year and month
  mutate(start = if_else(row_number() == 1, as.Date(paste(year, month, "01", sep = "-")), start)) %>% # Change date to first of the month for the first row in each group
  ungroup()
gt <- gt[, !names(gt) %in% c("year", "month")]

head(gt)
```


```{r}
CSI <- read.csv("/Users/kenjiwagner/Desktop/eecs448/projectstuff/EECS448Monthly.csv")
CSI <- CSI[c("DATE", "CSI")]
CSI$DATE <- as.Date(CSI$DATE)
head(CSI)
```

```{r}
# Merge Google Trends with CSI

trends <- merge(gt, CSI, by.x = "start", by.y = "DATE", all.x = TRUE)

trends
```

Plot data

```{r}
variables <- setdiff(colnames(trends), c("start", "CSI"))

trends_cp <- trends
trends_cp$CSI <- trends$CSI / 500

for (var in variables) {
  df <- melt(trends_cp[,c("start", "CSI", var)], id.vars = "start")
  df$value <- zoo::na.approx(df$value, na.rm = FALSE)

  g <- ggplot(df, aes(x = start, y = value, color = variable)) +
    geom_line() +
    labs(title = paste(var, " and CSI (scaled) over time"), x = "Date", y = "Value") +
    theme_minimal()
  print(g)
}
```

Ignore lagged data and only work with google trends at time of CSI
```{r}
#trends_interp <- trends
#trends_interp$CSI <- zoo::na.approx(trends_interp$CSI, na.rm = FALSE)
trends_interp <- na.omit(trends)
```

Pearson correlation of each variable with CSI
```{r}
cor_matrix <- cor(trends_interp[, -which(names(trends_interp) == "start")], use = "complete.obs")
cor_matrix["CSI",]
```

Recession, inflation, housing_market, used_car, gas_price, and layoff all have moderate correlation (|R| > 0.3)

Plotting CSI and variables on same plot with two y-axis (can be misleading, but useful to see how they move together)
```{r}

CSI <- xyplot(CSI~ start,
               data = trends_interp, type = "l")
var <- xyplot(used_car~ start,
              data = trends_interp, type = "l")
latticeExtra::doubleYScale(CSI,var,style1 = 5, style2 = 3, add.ylab2 = TRUE)
```

Test stationarity of CSI
```{r}
adf_result = adf.test(trends_interp$CSI, alternative = "stationary")
print(adf_result)
```

Cross Correlation of each variable with CSI
```{r}
variables <- setdiff(colnames(trends), c("start", "CSI"))

trends_cp <- trends_interp

lags_max <- 30

for (var in variables) {
  ccf_result <- ccf(trends_cp[[var]], trends_cp$CSI, plot = FALSE, lag.max = lags_max)

  correlations <- ccf_result$acf
  lags <- seq(-lags_max, lags_max)
  
  plot(lags, correlations, type = "h", ylim = c(-1, 1), xlab = "Lag", ylab = "Cross-Correlation",
       main = paste(var, ", CSI cross-correlation"))
    abline(h = 0, col = "black")
    abline(h = 0.06, col = "blue")
    abline(h = -0.06, col = "blue")
}
```

Note: positive lag means variable precedes CSI



Function to run Granger Causality Test over multiple windows to determine causality between variables and CSI
```{r}
library(vars)
library(zoo)


rolling_granger_test <- function(trends_interp, window_size = 156, step_size = 52) {
  
  results <- list()
  
  # Loop through variables
  for (var in names(trends_interp)) {
    if (var %in% c("CSI", "start") || !is.numeric(trends_interp[[var]])) next
    var_results <- list()
    
    start <- 1
    while (start < (nrow(trends_interp) - window_size + 1)) {
      end <- start + window_size - 1
      
      # Extract the window of data for CSI and the current variable and convert to time series
      window_data <- trends_interp[start:end, c("CSI", var)]
      if (any(is.na(window_data))) next
      window_ts <- ts(window_data, frequency = 12)
      
      # Perform Granger causality test and store p-value
      try({
        test_result <- grangertest(window_data[, var], window_data[, "CSI"], order = 1)
        var_results[[paste0("Start:", start, " End:", end)]] <- test_result$"Pr(>F)"[2]
      })
      
      start <- start + step_size
    }
    
    results[[var]] <- var_results
  }
  
  # print mean and standard deviation of each p-value
  stats <- lapply(results, function(p_values) {
    numeric_p_values <- as.numeric(p_values)
    
    list(
      mean = mean(numeric_p_values, na.rm = TRUE),
      sd = sd(numeric_p_values, na.rm = TRUE)
    )
  })
  
  for (var in names(stats)) {
    cat("Variable:", var, "\n")
    
    mean_p_value <- stats[[var]]$mean
    sd_p_value <- stats[[var]]$sd
    
    # Return number of asterisks based on significance level
    significance_stars <- function(p_value) {
      if (p_value < 0.01) {
        return("***")
      } else if (p_value < 0.05) {
        return("**")
      } else if (p_value < 0.1) {
        return("*")
      } else {
        return("")
      }
    }
    
    mean_stars <- significance_stars(mean_p_value)
    sd_stars <- significance_stars(sd_p_value)
    
    cat("Mean of p-values:", mean_p_value, mean_stars, "\n")
    cat("Standard deviation of p-values:", sd_p_value, sd_stars, "\n\n")
  }
}

```

Rolling Granger test of each variable with CSI
```{r}
rolling_granger_test(trends_interp)
```


Explore causal relationship between predictor variables and CSI when predictor variables are lagged. Change n below to explore different lags.
```{r}
trends_lagged <- trends %>%
  mutate(across(-c(start, CSI), lag, n = 4)) %>%
  na.omit()

rolling_granger_test(trends_lagged)

```


Recalculate CCF and cointegration with predictor variables lagged

